Food-web structure, stochasticity and response diversity drive temporal stability in complex communities

### 2023-10-11 _targets.Rmd
New plots

```{targets p_sem_tot_stab, tar_simple=TRUE}
tar_read(sem_eff_tab) %>%
  filter(effect_type == "total") %>%
  filter(response %in% c("async", "pop_stab", "stab_com")) %>%
  filter(!predictor %in% c("evenness_sae", "sae_even", "cpe_int", "cpe_env", "pop_stab", "async")) %>%
  mutate(across(c(response, predictor), ~var_replacement()[.x])) %>%
  ggplot(aes(
      x = predictor,
      y = effect,
      ymin = lower_ci,
      ymax = upper_ci)
    ) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(cols = vars(response)) +
  labs(x = "Predictor", y = "Total effect") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

@@

mod_tot <- lm(stab_com ~
  richness  + rho + env_stoch +
  ct_alive + w_avg_tlvl + Z +
  avg_int_strength +
  h +
  richness * rho +
  richness * env_stoch +
  richness * ct_alive +
  richness * avg_int_strength +
  richness * h +
  richness * w_avg_tlvl +
  richness * Z +
  richness * rho * ct_alive +
  richness * rho * w_avg_tlvl +
  richness * rho * avg_int_strength +
  richness * rho * h +
  richness * rho * Z +
  richness * env_stoch * ct_alive +
  richness * env_stoch * Z +
  richness * env_stoch * avg_int_strength+
  richness * env_stoch * h +
  richness * env_stoch * w_avg_tlvl,
  mutate(sim_fw, across(c(stab_com, richness), log)) #, richness
  )
check_collinearity(lm(stab_com ~ richness + rho +
    env_stoch + avg_int_strength +
  ct_alive + w_avg_tlvl + Z + h, sim_fw))
plot(mod_tot)

pred_data <- list(
  richness = log(seq(1, 30, 1)),
  ct_alive = c(.05, .10, .30),
  rho = c(0, .75, 1),
  avg_int_strength = c(.01, .02, .05, .1),
  env_stoch = c(.1, .3),
  w_avg_tlvl = c(1.5, 2, 3),
  Z = c(10, 100),
  h = c(2, 3)
  ) %>% expand.grid()
pred_data$stab_com <- predict(mod_tot, newdata = pred_data)
```

```{r}
p_pred_sr_int <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    rho %in% c(0, .75),
    Z == 10,
    h %in% c(3),
    avg_int_strength %in% c(.01, .05),
    w_avg_tlvl == 1.5
    ) %>%
  ggplot(aes(
      x = richness, y = stab_com,
      color = as.factor(rho),
      linetype = as.factor(avg_int_strength),
      )) +
  geom_line(linewidth = 2) +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Community stability (Log)",
    linetype = "Avg. Interaction strength",
    color = "Environmental correlation"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
ggsave(
  here("report", "figures", "p_pred_sr_int.png"),
  plot = p_pred_sr_int,
  scale = 2.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```

```{r}
plot_grid(
  tar_read(p_stab_rich_env_rho) + theme(legend.position = "none"),
  p_pred_sr_int)
```

```{r}

p_pred_sr_avg_tlvl_h <- pred_data %>%
  filter(
    ct_alive %in% c(.3),
    h %in% c(2,3),
    rho %in% c(0, 1),
    Z == 10,
    avg_int_strength %in% c(.1),
    w_avg_tlvl == c(1.5, 3)) %>%
  ggplot(aes(x = richness, y = stab_com, linewidth = as.factor(w_avg_tlvl), color = as.factor(h))) +
  geom_line() +
  facet_grid(cols = vars(env_stoch), rows = vars(rho), labeller = label_both) +
  theme_bw()
ggsave(
  here("report", "figures", "p_pred_sr_avg_tlvl_h.png"),
  plot = p_pred_sr_avg_tlvl,
  scale = 2.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```

>>>>

```{targets p_sem_tot_stab, tar_simple=TRUE}
tar_read(sem_eff_tab) %>%
  filter(effect_type == "total") %>%
  filter(response %in% c("async", "pop_stab", "stab_com")) %>%
  filter(!predictor %in% c("evenness_sae", "sae_even", "cpe_int", "cpe_env", "pop_stab", "async", "S", "ct")) %>%
  mutate(across(c(response, predictor), ~var_replacement()[.x])) %>%
  ggplot(aes(
      x = predictor,
      y = effect,
      ymin = lower_ci,
      ymax = upper_ci)
    ) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(cols = vars(response)) +
  labs(x = "Predictor", y = "Total effect") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

@@

mod_tot <- lm(stab_com ~
  richness  + rho + env_stoch +
  ct_alive + w_avg_tlvl + Z +
  avg_int_strength +
  h +
  richness * rho +
  richness * env_stoch +
  richness * ct_alive +
  richness * avg_int_strength +
  richness * h +
  richness * w_avg_tlvl +
  richness * Z +
  richness * rho * ct_alive +
  richness * rho * w_avg_tlvl +
  richness * rho * avg_int_strength +
  richness * rho * h +
  richness * rho * Z +
  richness * env_stoch * ct_alive +
  richness * env_stoch * Z +
  richness * env_stoch * avg_int_strength+
  richness * env_stoch * h +
  richness * env_stoch * w_avg_tlvl,
  mutate(sim_fw, across(c(stab_com, richness), log)) #, richness
  )
mod_tot_async <- lm(async ~
  richness  + rho + env_stoch +
  ct_alive + w_avg_tlvl + Z +
  avg_int_strength +
  h +
  richness * rho +
  richness * env_stoch +
  richness * ct_alive +
  richness * avg_int_strength +
  richness * h +
  richness * w_avg_tlvl +
  richness * Z +
  richness * rho * ct_alive +
  richness * rho * w_avg_tlvl +
  richness * rho * avg_int_strength +
  richness * rho * h +
  richness * rho * Z +
  richness * env_stoch * ct_alive +
  richness * env_stoch * Z +
  richness * env_stoch * avg_int_strength+
  richness * env_stoch * h +
  richness * env_stoch * w_avg_tlvl,
  mutate(sim_fw, across(c(async, richness), log)) #, richness
  )
mod_tot_pop_stab <- lm(pop_stab ~
  richness  + rho + env_stoch +
  ct_alive + w_avg_tlvl + Z +
  avg_int_strength +
  h +
  richness * rho +
  richness * env_stoch +
  richness * ct_alive +
  richness * avg_int_strength +
  richness * h +
  richness * w_avg_tlvl +
  richness * Z +
  richness * rho * ct_alive +
  richness * rho * w_avg_tlvl +
  richness * rho * avg_int_strength +
  richness * rho * h +
  richness * rho * Z +
  richness * env_stoch * ct_alive +
  richness * env_stoch * Z +
  richness * env_stoch * avg_int_strength+
  richness * env_stoch * h +
  richness * env_stoch * w_avg_tlvl,
  mutate(sim_fw, across(c(pop_stab, richness), log)) #, richness
  )
check_collinearity(lm(stab_com ~ richness + rho +
    env_stoch + avg_int_strength +
  ct_alive + w_avg_tlvl + Z + h, sim_fw))
plot(mod_tot)

pred_data <- list(
  richness = log(seq(1, 30, 1)),
  ct_alive = c(.05, .10, .30),
  rho = c(0, .75, 1),
  avg_int_strength = c(.01, .02, .05, .1),
  env_stoch = c(.1, .6),
  w_avg_tlvl = c(1.5, 2, 3),
  Z = c(10, 100),
  h = c(2, 3)
  ) %>% expand.grid()
pred_data$stab_com <- predict(mod_tot, newdata = pred_data)
pred_data$async <- predict(mod_tot_async, newdata = pred_data)
pred_data$pop_stab <- predict(mod_tot_pop_stab, newdata = pred_data)
pred_data$resp_div <- rho_replacement()[as.character(pred_data$rho)]
```

```{r}
p_pred_sr_int <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    rho %in% c(0, .75),
    richness >= 1,
    Z == 10,
    h %in% c(3),
    avg_int_strength %in% c(.01, .1),
    w_avg_tlvl == 1.5
    ) %>%
  ggplot(aes(
      x = richness, y = stab_com,
      color = as.factor(resp_div),
      linetype = as.factor(avg_int_strength),
      alpha = as.factor(avg_int_strength)
      )) +
  ylim(2, 8) +
  geom_line(linewidth = 1) +
   scale_alpha_manual(values=c(1, 0), guide = "none") +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Community stability (Log)",
    linetype = "Avg. Interaction strength",
    color = "Response diversity"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
p_pred_sr_int1 <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    rho %in% c(0, .75),
    Z == 10,
    h %in% c(3),
    avg_int_strength %in% c(.01, .1),
    w_avg_tlvl == 1.5
    ) %>%
  ggplot(aes(
      x = richness, y = stab_com,
      color = as.factor(resp_div),
      linetype = as.factor(avg_int_strength),
      alpha = as.factor(avg_int_strength)
      )) +
  geom_line(linewidth = 1) +
   scale_alpha_manual(values=c(1, 0), guide = "none") +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Community stability (Log)",
    linetype = "Avg. Interaction strength",
    color = "Response diversity"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
p_pred_sr_int2 <- p_pred_sr_int1 +
   scale_alpha_manual(values=c(1, 1), guide = "none")
ggsave(
  here("report", "figures", "p_pred_sr_int1.png"),
  plot = p_pred_sr_int1,
  scale = 2.4,
  width = 88,
  height = 88 * .5,
  units = "mm"
)
ggsave(
  here("report", "figures", "p_pred_sr_int2.png"),
  plot = p_pred_sr_int2,
  scale = 2.4,
  width = 88,
  height = 88 * .5,
  units = "mm"
)
```

```{r}
pred_data_show <- pred_data %>%
  filter(
    ct_alive %in% c(.3),
    rho %in% c(0, .75),
    Z == 100,
    h %in% c(3),
    avg_int_strength %in% c(.01, .1),
    w_avg_tlvl == 1.5
    )

sim_fw %>%
  filter(env_stoch %in% c(.1, .6), rho %in% c(0, .75), h == 3, Z == 100) %>%
  ggplot(aes(x = log(richness), y = log(stab_com), color = as.factor(rho))) +
  geom_point(alpha = .3) +
  geom_line(data = pred_data_show,
    aes(x = richness, y = stab_com,
      color = as.factor(rho),
      linetype = as.factor(avg_int_strength),
      alpha = as.factor(avg_int_strength))
    ) +
  scale_alpha_manual(values=c(1, 0), guide = "none") +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Community stability (Log)",
    color = "Environmental correlation"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
leg_pred <- get_legend(p_pred_sr_int)
p_sim_pred <- plot_grid(
  plot_grid(
    tar_read(p_stab_rich_env_rho) + theme(legend.position = "none"),
    p_pred_sr_int + theme(legend.position = "none"),
  labels = "auto"
  ),
  leg_pred,
  nrow = 2,
  rel_heights = c(1, .1)
)
ggsave(
  here("report", "figures", "p_sim_pred.pdf"),
  plot = p_sim_pred,
  scale = 2.4,
  width = 88,
  height = 88 * .4,
  units = "mm"
)
```

```{r}
p_pred_sr_int_async <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    rho %in% c(0, .75),
    Z == 10,
    h %in% c(3),
    avg_int_strength %in% c(.01, .1),
    w_avg_tlvl == 1.5
    ) %>%
  ggplot(aes(
      x = richness, y = async,
      color = as.factor(rho),
      linetype = as.factor(avg_int_strength),
      )) +
  geom_line(linewidth = 1) +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Asynchrony (Log)",
    linetype = "Avg. Interaction strength",
    color = "Environmental correlation"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
ggsave(
  here("report", "figures", "p_pred_sr_int_async.png"),
  plot = p_pred_sr_int_async,
  scale = 2.4,
  width = 88,
  height = 88 * .5,
  units = "mm"
)
```

```{r}
p_pred_sr_int_pop_stab <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    rho %in% c(0, .75),
    Z == 10,
    h %in% c(3),
    avg_int_strength %in% c(.01, .1),
    w_avg_tlvl == 1.5
    ) %>%
  ggplot(aes(
      x = richness, y = pop_stab,
      color = as.factor(rho),
      linetype = as.factor(avg_int_strength),
      )) +
  geom_line(linewidth = 1) +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Population stability (Log)",
    linetype = "Avg. Interaction strength",
    color = "Environmental correlation"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
ggsave(
  here("report", "figures", "p_pred_sr_int_pop_stab.png"),
  plot = p_pred_sr_int_pop_stab,
  scale = 2.4,
  width = 88,
  height = 88 * .5,
  units = "mm"
)
```

```{r}
p_pred_sr_avg_tlvl_h <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    h == c(3),
    rho %in% c(0, .75),
    Z == 10,
    avg_int_strength %in% c(.1),
    w_avg_tlvl == c(1.5, 3)) %>%
  ggplot(aes(x = richness, y = stab_com,
      linetype = as.factor(w_avg_tlvl), color = as.factor(rho))) +
  geom_line() +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  theme_bw()
ggsave(
  here("report", "figures", "p_pred_sr_avg_tlvl_h.png"),
  plot = p_pred_sr_avg_tlvl,
  scale = 2.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```

### 2023-10-11 paper/manuscript.Rmd
New plots

---
title: Food-web structure, stochasticity and response diversity drive temporal stability in complex communities
author: Alain Danet, Andrew P. Beckerman
output:
  bookdown::pdf_document2
always_allow_html: true
toc: false
bibliography: references.bib
---

@@

@@

Biodiversity, in terms of species richness, has been repeatedly shown to drive
higher temporal stability of ecological communities, across various taxa and
ecosystems. Species richness can dampen the effects of perturbations on the
temporal variability of community biomass because more diverse communities
contain species with more diversity of responses to perturbations, which
generates asynchrony in population fluctuations, thereby stabilizing community
biomass fluctuations [@loreau_species_2008; @thibaut_understanding_2013;
@zhao_biodiversity_2022]. In turn, the effects of human pressures on biotic
homogenization reported in various ecosystems [@buisson_toward_2013;
@muthukrishnan_invasive_2020; @saladin_rapid_2020] is another threat, which
might in turn affect dampens the effects of species richness on ecosystem
stability. However, the joint effects of external stressors such as
environmental variabilityand internal stressors such as the
loss response diversity has been rarely assessed in temporal stability areas.

@@

```{r, fig.cap=concept_cap}
knitr::include_graphics(here("report", "figures", "conceptual_figures.pdf"))
```

>>>>

---
title: Food-web structure, stochasticity and response diversity drive temporal stability in complex communities
author: Alain Danet, Andrew P. Beckerman
output:
  bookdown::pdf_document2
always_allow_html: true
toc: false
bibliography: references.bib
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
---

@@

# IDEA, meeting with Sonia

- Simplifying SEM in main text, move details to appendix
- State diagram for stability - richness relationships along two food-web
  metrics and panels varying response diversity.
- Look at the french data for portefolio and compensatory dynamics effects.
- Mid-October

# Abstract

Current biotic homogenization and increasing frequency of extreme events are
projected to dramatically destabilize ecological communities. Understanding how
complex communities such as food-webs respond to such threats is crucial to take
adequate conservation measures. We then investigated how response diversity,
environmental stochasticity, and food-web structure drive temporal stability of
biomass in complex food-webs, using a stochastic bioenergetic food-web model. We
found that higher average trophic level, omnivory degree and generalism
increased  average population stability while species richness had little
effects. At the opposite higher average trophic level, omnivory degree and
generalism decreased asynchrony and compensatory dynamics raising trophic
interactions. We further found that that high interaction strength drive
negative stability-species richness when response diversity is low but that high
response diversity always drive positive stability-richness relationship. Our
study reveals that response diversity and food-web structure drive temporal
stability and further improve our understanding of biodiversity-stability
relationships in complex trophic communities.

@@

Biodiversity, in terms of species richness, has been repeatedly shown to drive
higher temporal stability of ecological communities, across various taxa and
ecosystems. Species richness can dampen the effects of perturbations on the
temporal variability of community biomass because more diverse communities
contain species with more diversity of responses to perturbations, which
generates asynchrony in population fluctuations, thereby stabilizing community
biomass fluctuations [@loreau_species_2008; @thibaut_understanding_2013;
@zhao_biodiversity_2022]. In turn, the effects of human pressures on biotic
homogenization reported in various ecosystems [@buisson_toward_2013;
@muthukrishnan_invasive_2020; @saladin_rapid_2020] is another threat, which
might in turn affect dampens the effects of species richness on ecosystem
stability. However, the joint effects of external stressors such as
environmental variability and internal stressors such as the
loss response diversity has been rarely assessed in temporal stability areas.

@@

```{r, eval = FALSE, fig.cap=concept_cap}
knitr::include_graphics(here("report", "figures", "conceptual_figures.pdf"))
```
